---
title: "DATA 605 Assignment 1"
author: "Ahsanul Choudhury"
date: "February 5, 2017"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Problem set 1

   (1) Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; -4]
   
   **Sol:**

```{r vector, echo=TRUE}
u <- c(0.5, 0.5)
v <- c(3, -4)
```

     The dot product is `r u%*%v`.


   (2) What are the lengths of u and v? Please note that the mathematical notion of the
length of a vector is not the same as a computer science definition.

   **Sol:**
   
    The length of u is `r sqrt(u%*%u)` and length of v is `r sqrt(v%*%v)`.


   (3) What is the linear combination: 3u - 2v?

   **Sol:**
   
    The linear combination is `r 3*u - 2*v`.

   (4) What is the angle between u and v

   **Sol:**
    The angle between u and v is `r acos(u %*% v / (sqrt(u %*% u) * sqrt(v %*% v)))`

***

###Problem set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. You don't have to worry about degenerate
cases and can safely assume that the function will only be tested with a system of equations
that has a solution. Please note that you do have to worry about zero pivots, though. Please
note that you should not use the built-in function solve to solve this system or use matrix
inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.

Please test it with the system below and it should produce a solution x = [-1.55, -0.32, 0.95]

$$
\begin{bmatrix}
1 & 1 & 3 \\
2 & -1 & 5 \\
-1 & -2 & 4
\end{bmatrix}\,
\begin{bmatrix}
x_1 \\x_2 \\x_3
\end{bmatrix}=
\begin{bmatrix}
1 \\2 \\6
\end{bmatrix}
$$

###Solution to problem set 2

Create Function

```{r function, echo=TRUE}

sol <- function(A, b){
 #Pivot
  pivot <- A[1,1]
  #exchange rows where pivot = 0
  if (pivot == 0){
    A <- A[c(2,1,3),]
  pivot <- A[1,1]
  if (pivot == 0){
    A <- A[c(3,2,1),]
  pivot <- A[1,1]
  }
  }
  #Multiplier
  mplyr_1 <- A[2,1]/pivot
  mplyr_2 <- A[3,1]/pivot
  #Elimination: row2
  A[2, ] <- A[2, ] - mplyr_1 * A[1, ]
  b[2,] <- b[2,] - mplyr_1 * b[1,]
  #Elimination: row3
    A[3, ] <- A[3, ] - mplyr_2 * A[1, ]
  b[3,] <- b[3,] - mplyr_2 * b[1,]
  
  #Repeat
  
  pivot_1 <- A[2,2]
  #exchange rows where pivot = 0
  if (pivot_1 == 0){
    A <- A[c(1,3,2),]
  pivot_1 <- A[2,2]
  }
  #Multiplier
  mplyr_3 <- A[3,2]/pivot_1
  #Elimination: row3
  A[3, ] <- A[3, ] - mplyr_3 * A[2, ]
  b[3,] <- b[3,] - mplyr_3 * b[2,]
  
  #x3 solution
  x3 <- b[3] / A[3, 3]
  
  #substitution1
  x2 <- (b[2] - A[2, 3] * x3) / A[2, 2]
  #substitution2
  x1 <- (b[1] - A[1, 3] * x3 - A[1, 2] * x2) / A[1, 1]
  
  #results
  x <- matrix(c(x1, x2, x3), nrow = 3)
  x
  
}
```

Test Function

```{r test, echo=TRUE}
A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim = c(3,3))
b <- array(c(1, 2, 6), dim = c(3,1)) 
round(sol(A, b), 2)
```

***

