---
title: "DATA 605 Assignment 2"
author: "Ahsanul Choudhury"
date: "February 12, 2017"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Problem set 1

(1) Show that $A^TA\neq AA^T$ in general. (Proof and demonstration.)

**Sol:**

Let A = $(m \times n)$ =
$$
{\begin{bmatrix}
 x_{1} \; x_{2} \; x_{3} \\ 
 y_{1} \; y_{2} \; y_{3} \\ 
 z_{1} \; z_{2} \; z_{3} 
\end{bmatrix}}
$$
$\therefore  A^T = (n \times m)$ = 
$$
{\begin{bmatrix}
 x_{1} \; y_{1} \; z_{1} \\ 
 x_{2} \; y_{2} \; z_{2} \\ 
 x_{3} \; y_{3} \; z_{3} 
\end{bmatrix}}
$$
$A^TA$ =

$$
{\begin{bmatrix}
 {x_{1}}^{2} + {y_{1}}^{2} + {z_{1}}^{2} \;\;\;\;\;\;\;\;\;\;\;\;\; x_{1} x_{2} + y_{1} y_{2} + z_{1} z_{2} \;\;\;\;\;\;\; x_{1} x_{3} + y_{1} y_{3} + z_{1} z_{3} \\ 
 x_{2} x_{1} + y_{2} y_{1} + z_{2} z_{1} \;\;\;\;\;\;\; {x_{2}}^{2} + {y_{2}}^{2} + {z_{2}}^{2} \;\;\;\;\;\;\;\;\;\;\;\; x_{2} x_{3} + y_{2} y_{3} + z_{2} z_{3} \\ 
 x_{3} x_{1} + y_{3} y_{1} + z_{3} z_{1} \;\;\;\;\;\;\;\; x_{3} x_{2} + y_{3} y_{2} + z_{3} z_{2} \;\;\;\;\;\;\;\;\;\;\;\; {x_{3}}^{2} + {y_{3}}^{2} + {z_{3}}^{2} \\ 
\end{bmatrix}}
$$

$AA^T$ =
$$
{\begin{bmatrix}
 {x_{1}}^{2} + {x_{2}}^{2} + {x_{3}}^{2} \;\;\;\;\;\;\;\;\;\;\;\; x_{1} y_{1} + x_{2} y_{2} + x_{3} y_{3} \;\;\;\;\;\;\; x_{1} z_{1} + x_{2} z_{2} + x_{3} z_{3} \\ 
 y_{1} x_{1} + y_{2} x_{2} + y_{3} x_{3} \;\;\;\;\;\;\; {y_{1}}^{2} + {y_{2}}^{2} + {y_{3}}^{2} \;\;\;\;\;\;\;\;\;\;\;\;\;\; y_{1} z_{1} + y_{2} z_{2} + y_{3} z_{3} \\ 
 z_{1} x_{1} + z_{2} x_{2} + z_{3} x_{3} \;\;\;\;\;\;\;\; z_{1} y_{1} + z_{2} y_{2} + z_{3} y_{3} \;\;\;\;\;\;\;\;\;\;\;\;\;\; {z_{1}}^{2} + {z_{2}}^{2} + {z_{3}}^{2} \\ 
\end{bmatrix}}
$$
Example:

A=
```{r matrix_A, echo=TRUE}
A <- matrix(seq(from=1,to=9), nrow=3)
A
```

$A^T$=
```{r matrix_A_T, echo=TRUE}
A_T <- t(A)
A_T
```


$A^TA$ =
```{r ATA, echo=TRUE}
t(A)%*%A
```

$AA^T$ =
```{r AAT, echo=TRUE}
A%*%t(A)
```


$A^TA\neq AA^T$ in general.



(2) For a special type of square matrix A, we get AT $A^TA= AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

**Answer:**

$A^TA= AA^T$ only occurs when $A^T$ = A. In other words, the transpose of the matrix is the same as the matrix itself. 


***


###Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

**Sol:**

```{r function_LDU, echo=TRUE}
fun_LU <- function(A){
  #Check for valid input
  if(nrow(A) != ncol(A)){
    stop("Input must be square matrix.")
  }
    else {
    n <- nrow(A)
    L <- diag(n)
    U <- A
      for (i in 1:(n - 1)) {
      for (j in (i + 1):n) {
        # get multipliers
        L[j, i] <- U[j, i] / U[i, i]
        # pivots and multiplication
        U[j, ]  <- U[j, ] - L[j, i] * U[i, ]
      }
    }
    
        #Get results
        LU <- list("L" = L, "U" = U)
    }
}

```   


**Testing the function using Matrix given in lecture note for week 2**
```{r find_LU, echo=TRUE}
A <- matrix(c(1,2,3,1,1,1,2,0,1),nrow=3)
A
LU <- fun_LU(A)
LU
```


```{r test, echo=TRUE}
fun_LU(A)$L %*% fun_LU(A)$U 

fun_LU(A)$L %*% fun_LU(A)$U  == A
```

