---
title: "DATA 605 Assignment 13"
author: "Ahsanul Choudhury"
date: "May 7, 2017"
output:
  html_document:
     theme: cerulean
     toc: true
     toc_float:
       collapsed: false
       smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


-   Write a program to compute the derivative of $f(x) = x^3 + 2x^2$ at any value of $x$. Your function should take in a value of $x$ and return back an approximation to the derivative of $f(x)$ evaluated at that value. You should not use the analytical form of the derivative to compute it. Instead, you should compute this approximation using limits.


-   Now, write a program to compute the area under the curve for the function $3x^2+4x$ in the range $x = [1, 3]$. You should first split the range into many small intervals using some really small $???x$ value (say $1e-6$) and then compute the approximation to the area under the curve.


Please solve these problems analytically (i.e. by working out the math) and submit your answers.

-   Use integration by parts to solve for $\int sin(x)cos(x)dx$

-   Use integration by parts to solve for $\int x^2e^xdx$

-   What is $\frac{d}{dx}(x cos(x))$?

-   What is $\frac{d}{dx}({e^x}^4)$?


##Problem set 1

### 1.1 Compute Derivative

```{r driv, echo=TRUE}
my_drivative <- function(x){
  driv <- ((x+1e-5)^3 + 2*(x+1e-5)^2 - x^3 - 2*x^2)/1e-5
  driv <- round(driv,0)
  return(driv)
}

my_drivative(1)
my_drivative(5)
my_drivative(10)
```

### 1.2 Area Under the Curve

```{r area, echo=TRUE}
# split the range into many small intervals
x <- seq(1, 3, by=1e-6)
# function
f <- 3*(x^2) + 4*x
area <- f*(1e-6)
sum(area)
```

##Problem set 2

###2.1

$\int sin(x)cos(x)dx$

let, $u = cos(x)$

  $du = -sin(x)dx$
     
  $dx = \frac{du}{-sin(x)}$
     
  $\int sin(x) u \frac{du}{-sinx}$
     
  $=-\frac{1}{2}udu$
     
  $=-\frac{1}{2}u^2$
     
  $=-\frac{1}{2}cos^2x + c$
  
  
###2.2

$\int x^2e^xdx$

We know, $\int fg'dx = fg - \int f'gdx$

If $f = x^2$, then, $f' = 2x$, if $g = e^x$ then $g' = e^x$

$\int x^2e^xdx$

$= x^2e^x - 2\int xe^xdx$

(we will perform another integration by parts for $\int xe^xdx$ where $f(x) = x, f'(x) = 1, g(x) = e^x. g'(x) = e^x$)

$= x^2e^x - 2(xe^x - \int e^xdx)$

$= x^2e^x - 2(xe^x-e^x) + C$

$= x^2e^x-2xe^x + 2e^x + C$


###2.3

$\frac{d}{dx}(xcos(x))$

$= x(-sin(x)) + cos(x)(1)$

$= - xsin(x) + cos(x)$


###2.4

$\frac{d}{dx}(e^{x^{4}})$

$= e^{x^{4}}.\frac{d}{dx}[x^{4}]$

$4x^3{e^x}^4$