---
title: "DATA 605 Assignment 5"
author: "Ahsanul Choudhury"
date: "March 5, 2017"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Problem set 1

   In this problem set we'll work out some properties of the least squares solution that wereviewed in the weekly readings. Consider the unsolvable system Ax = b as given below:

$$
\begin{bmatrix}
1 & 0 \\
1 & 1 \\
1 & 3 \\
1 & 4
\end{bmatrix}
\begin{bmatrix}
x_1 \\
x_2
\end{bmatrix}
=
\begin{bmatrix}
0 \\
8 \\
8 \\
20
\end{bmatrix}
$$

   - Write R markdown script to compute $A^TA$ and $A^Tb$.
   
   ***Sol**
   
```{r ATA/b, echo=TRUE}
A <- matrix(c(1,0,1,1,1,3,1,4), byrow=TRUE, nrow=4)
b <- matrix(c(0,8,8,20), nrow=4)
(ATA <- t(A) %*% A)
(ATb <- t(A) %*% b)
```


   - Solve for $\hat{x}$ in R using the above two computed matrices.
   
   ***Sol***
   
```{r xhat, echo=TRUE}
(x_hat <- solve(ATA) %*% ATb)

# check work with lsfit()(sanity test)
x_hat_lsfit <- lsfit(A,b,intercept=FALSE)
x_hat_lsfit$coef
```

   - What is the squared error of this solution?
   
   ***Sol***
   
```{r sqerror, echo=TRUE}
p <- A %*% x_hat
# compute the error vector e = b - p
(e <- b-p)

# check work with lsfit()
round(e) == round(x_hat_lsfit$residuals)
# squared error
(sq_e <- sum(e^2))
```


   - Instead of $b = [0; 8; 8; 20]$, start with $p = [1; 5; 13; 17]$ and find the exact solution (i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector **e = 0**).
   
   ***Sol***
   
```{r p, echo=TRUE}
p <- matrix(c(1,5,13,17), nrow=4)
(ATp <- t(A) %*% p)

# solve x_hat using using p instead of b
(x_hat_p <- solve(ATA) %*% ATp)

# calculate the predicted values 
y_hat <- A %*% x_hat_p

# calculate error
(ep <- round(y_hat - p))
```


   - Show that the error $e = b - p = [-1; 3;-5; 3]$.
   
   ***Sol***
   
```{r error, echo=TRUE}
#residual error
(b-p)

#check with lsfit
(lsfit(A, b,intercept=FALSE)$residuals)
```


   - Show that the error **e** is orthogonal to **p** and to each of the columns of **A**.
   
```{r orthogonal, echo=TRUE}
# e*p == 0
(round(sum(e*p)))

#Show that e and column 1 of matrix A are orthogonal:
(round(sum(e*A[,1])))

#Show that e and column 2 of matrix A are orthogonal:
(round(sum(e*A[,2])))
```

***


###Problem set 2


   Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learningdataset). This dataset contains 5 columns: ***displacement, horsepower, weight, acceleration, mpg***. We are going to model ***mpg*** as a function of the other four variables.
   
   Write an R markdown script that takes in the auto-mpg data, extracts an **A** matrix from the first 4 columns and **b** vector from the fifth (***mpg***) column. Using the least squares approach, your code should compute the best fitting solution. That is, find the best fitting equation that expresses ***mpg*** in terms of the other 4 variables. Finally, calculate the fitting error between the ***predicted mpg*** of your model and the actual ***mpg***. Your script should be able to load in the 5 column data set, extract A and b, and perform the rest of the calculations. Please have adequate comments in your code to make it easy to follow your work.


   ***SOL***

```{r auto1, echo=TRUE}

# Setting working directory
# Note: Both .rmd and .data needs to be in same directory
set_directory <- function(directory) {
if (!is.null(directory))
   setwd(directory)
}

# Convert data into matrix format and give column name
auto_data<-as.matrix(read.table("auto-mpg.data", header=FALSE,as.is=TRUE))
colnames(auto_data) <- c("displacement", "horsepower","weight","acceleration","mpg")

# check data
head(auto_data)

# Extract an A matrix from the first four columns and vector b from the fifth.
A <- as.matrix(auto_data[,1:4])
b <- as.matrix(auto_data[,5])

# Adding a column of repeating ones for the Intercepts
A <- cbind(rep(1,nrow(A)),A)
```

   Using the least squares approach, find best fitting equation that expresses ***mpg*** in terms of the other 4 variables:
   
   $$\hat{mpg} = \hat{x}[1] + displacement *  \hat{x}[2] + horsepower *  \hat{x}[3] + weight *  \hat{x}[4] + acceleration *  \hat{x}[5]$$
   
```{r auto2, echo=TRUE}
ATA <- t(A) %*% A
ATb <- t(A) %*% b

# calculate x_hat for best fitting solution
(x_hat <- solve(ATA) %*% ATb)

# check it with lsfit()
(ls_check <- lsfit(A, b, intercept=FALSE)$coefficients)
```

   calculate the fitting error between the ***predicted mpg*** of your model and the actual ***mpg***.
   
```{r auto3, echo=TRUE}
p <- A %*% x_hat
e <- b-p
(sq_e <- sum(e^2))
```


